Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2020-07-21 manuscript/response_to_reviews.Rmd

327 I agree that this is the point that most deserves discussion in paragraph 2

**Thank you.**

334 You might go into detail regarding what you think this says about the causal hypothesis behind the global decline of large individuals. To me it helps reject notions that somehow larger trees are more exposed to wind throw or other damaging agents. It is consistent with the hypothesis that larger individuals have wider conduits, and (assuming that wider conduits are more vulnerable to drought induced embolism) thus larger individuals are (all else being equal…) more vulnerable to embolism than smaller conspecifics.

*KAT, come back to this*

336 beautifully said “impossible for trees to efficiently transport water to great heights and simultaneously maintain strong resistance and resilience to drought-induced embolism”

@@

**Thank you.**

349 fascinating; it seems plausible that current climate changed induced large-individual mortality involves trees growing to conduit diameters, and thus heights, permitted under previously more even and moister climates, and now finding themselves in more extreme climates that make their heights and conduits excessively wide given current conditions.

*https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/122*

@@

350 if you wanted to expand your discussion of causal mechanisms or the implications of your findings, I think you could jettison this paragraph. I admire your thoroughness in offering caveats, but I agree that this does not seem like a serious problem for your study.

**We have changed our treatment of canopy position and removed this paragraph.**

### 2020-07-21 manuscript/climate_sensitivity_draft.Rmd
worked on R2

Belowground, taller trees would tend to have larger root systems (*Enquist and Niklas 2001; DOI: 10.1126/science.1066360*), but this does not necessarily imply that they have greater access to or reliance on deep soil-water resources that may be critical during drought.
Rather, larger trees may allocate more to abundant shallow roots that are beneficial for taking up water from rainstorms (*Meinzer et al. 1999; DOI: 10.1007/s004420050931*).
In any case, the potentially greater access to water did not override the disadvantage conferred by height--and, in fact, greater moisture access in non-drought years (here, higher TWI) appears to make trees more sensitive to drought [@zuleta_drought-induced_2017; @stovall_tree_2019].
The observed height-sensitivity of $Rt$, together with the lack of conferred advantage to large stature in drier topographic positions, agrees with the concept that physiological limitations to transpiration under drought shift from soil water availability to the plant-atmosphere interface as forests age [@bretfeld_plant_2018], such that tall, dominant trees are the most sensitive in mature forests. 
Again, additional research comparing drought responses across forests with different tree heights and water availability would be valuable for disentangling the relative importance of above- and belowground mechanisms across trees fo different size.

@@

Concerted measurement of tree-rings and leaf hydraulic traits of emerging importance [@scoffoni_leaf_2014; @bartlett_correlations_2016; @medeiros_extensive_2019] allowed novel insights into the role of hydraulic traits in shaping drought response. The finding that $PLA_{dry}$ and $\pi_{tlp}$ can be useful for predicting drought responses of tree growth (Fig. **4**; Table 1) is both novel and consistent with previous studies linking these traits to habitat and drought tolerance. Previous studies have demonstrated that $\pi_{tlp}$ and $PLA_{dry}$ are physiologically meaningful traits linked to species distribution along moisture gradients [@marechaux_drought_2015; @fletcher_evolution_2018; @medeiros_extensive_2019; @simeone_coupled_2019;  @rosas_adjustments_2019], and our findings indicate that these traits also influence drought responses. Furthermore, the observed linkage of $\pi_{tlp}$ to $Rt$ in this forest aligns with observations in the Amazon that $\pi_{tlp}$ is higher in drought-intolerant than drought-tolerant plant functional type. Further, it adds support to the idea that this trait is useful for categorizing and representing species' drought responses in models [@powell_differences_2017]. Because both $PLA_{dry}$ and $\pi_{tlp}$ can be measured relatively easily [@bartlett_rapid_2012; @scoffoni_leaf_2014], they hold promise for predicting drought growth responses across diverse forests. The importance of predicting drought responses from species traits increases with tree species diversity; whereas it is feasible to study drought responses for all dominant species in most boreal and temperate forests (e.g., this study), this becomes difficult to impossible for species that do not form annual rings, and for diverse tropical forests. Although progress is being made for the tropics [@schongart_dendroecological_2017], a full linkage of hydraulic traits to drought responses would be invaluable for forecasting how little-known species and whole forests will respond to future droughts [@powell_differences_2017].

>>>>

Belowground, taller trees would tend to have larger root systems (*Enquist and Niklas 2001; DOI: 10.1126/science.1066360*), but this does not necessarily imply that they have greater access to or reliance on deep soil-water resources that may be critical during drought.
Rather, larger trees may allocate more to abundant shallow roots that are beneficial for taking up water from rainstorms (*Meinzer et al. 1999; DOI: 10.1007/s004420050931*).
In any case, the potentially greater access to water did not override the disadvantage conferred by height--and, in fact, greater moisture access in non-drought years (here, higher TWI) appears to make trees more sensitive to drought [@zuleta_drought-induced_2017; @stovall_tree_2019].
*This is likely because moister habitats would tend to support species and individuals with more mesophytic traits (e.g., higher $\pi_{tlp}$ and low $PLA_{dry}$, wider conduits), potentially growing to greater heights, and these are then more vulnerable when drought hits.*
The observed height-sensitivity of $Rt$, together with the lack of conferred advantage to large stature in drier topographic positions, agrees with the concept that physiological limitations to transpiration under drought shift from soil water availability to the plant-atmosphere interface as forests age [@bretfeld_plant_2018], such that tall, dominant trees are the most sensitive in mature forests. 
Again, additional research comparing drought responses across forests with different tree heights and water availability would be valuable for disentangling the relative importance of above- and belowground mechanisms across trees fo different size.

@@

Concerted measurement of tree-rings and leaf hydraulic traits of emerging importance [@scoffoni_leaf_2014; @bartlett_correlations_2016; @medeiros_extensive_2019] allowed novel insights into the role of hydraulic traits in shaping drought response. The finding that $PLA_{dry}$ and $\pi_{tlp}$ can be useful for predicting drought responses of tree growth (Fig. **4**; Table 1) is both novel and consistent with previous studies linking these traits to habitat and drought tolerance. Previous studies have demonstrated that $\pi_{tlp}$ and $PLA_{dry}$ are physiologically meaningful traits linked to species distribution along moisture gradients [@marechaux_drought_2015; @fletcher_evolution_2018; @medeiros_extensive_2019; @simeone_coupled_2019;  @rosas_adjustments_2019], and our findings indicate that these traits also influence drought responses. Furthermore, the observed linkage of $\pi_{tlp}$ to $Rt$ in this forest aligns with observations in the Amazon that $\pi_{tlp}$ is higher in drought-intolerant than drought-tolerant plant functional type. Further, it adds support to the idea that this trait is useful for categorizing and representing species' drought responses in models [@powell_differences_2017]. Because both $PLA_{dry}$ and $\pi_{tlp}$ can be measured relatively easily [@bartlett_rapid_2012; @scoffoni_leaf_2014], they hold promise for predicting drought growth responses across diverse forests. The importance of predicting drought responses from species traits increases with tree species diversity; whereas it is feasible to study drought responses for all dominant species in most boreal and temperate forests (e.g., this study), this becomes difficult to impossible for species that do not form annual rings, and for diverse tropical forests. Although progress is being made for the tropics [@schongart_dendroecological_2017], a full linkage of hydraulic traits to drought responses would be invaluable for forecasting how little-known species and whole forests will respond to future droughts ([Christoffersen et al. 2016](https://pearl.plymouth.ac.uk/handle/10026.1/12918)) [@powell_differences_2017].
